Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1123 at $z$=0.654

### 2016-05-24 WRITEUP/ms.tex
SED section close to done

\citestyle{aa}
\shorttitle{Something goes here}
\shortauthors{Leung \& Riechers}

@@

% moment 0 map: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth, first and second moment maps shown in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube, over $\Delta v$$\sim$750 \kms.
The first and second moment maps are clipped at 3$\sigma$.
The peak flux density is 8.12 $\pm$ 0.30 Jy\,\kms\,beam\pmOne in the intensity-integrated
map. The deconvolved source size is estimated to be 5\farcs1$\pm$0\farcs72 $\times$ 3\farcs72$\pm$0\farcs66
and thus the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as 2D-Gaussian,
the fit recovers the line intensity enclosed within the emitting
region, we therefore take this as an estimate on the extent of the lensed
emission. On the other hand, if we assume the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein Ring" and
the lensed knots (see \Fig{CO21mom}).

@@

The emission of the red component coincides with the Einstein Ring seen in the
optical image, with most of its apparent flux originating from the lensed arc
in the Southeast, whereas the blue component is predominately coming from
solely the lensed arc shown in \Fig{CO21mom}. To illustrate this, we show the
channel maps of 21.5\,\kms width and spatial spectra of 1\farcs5 resolution in
\Fig{chanmap} and \Fig{spatialSpec}, respectively. The figures show that emission
is present to the West and peaks toward the lensing arc (black across in \Fig{chanmap})
in the red wing, and shifts to the East with decreasing velocity (blue wing).
This indicates extended emission in the source plane, where
emission from different kinematic components is separated by different amounts
relative to the caustics. One consequence of such is the observed asymmetric
line profile. A detailed analysis on differential lensing of \bco is presented in \Sec{differential}.

@@

\subsection{Origin of the \bco Emission} \label{sec:origin} %DONE
Evidently, the CO emission is co-spatial with the lensed optical emission in
the background AGN host galaxy.
The question is then whether the optically
faint companion also contributes to the CO flux and thus contains a
non-negligible amount of molecular gas.
Indeed, based on our lens model presented in \Sec{lensmodel}, we find it
highly likely that part of the CO emission originates from
the companion. We thus classify RXJ1131 as a ``wet-wet'' merger.
However, we cannot quantify this with a mass ratio given the
spatial resolution, nor can we rule out the possibility that
the ``companion'' is instead a gas component/tidal feature ripped out
from previous interaction.
In the subsequent sections, we will interpret
the this additional source component as gas emission
in a nearby galaxy (i.e. a ``wet-wet'' merger), as suggested by previous studies of RXJ1131.

@@

We detect PdBI 2\,mm continuum in \Fig{cont}. The integrated flux density is
1.2$\pm$0.2 mJy, with a peak flux $S_\nu$ = 800$\pm$88 \micron Jy\,beam\pmOne
centering on the lensing galaxy. A slightly extended emission is also seen
along the lensed arc. This suggests that the detected emission comes from
both the foreground galaxy and the background galaxy and that the
emission is marginally resolved along its major axis.
We subtract a point source model in $uv$-plane to remove the unresolved
emission toward the foreground galaxy. The peak flux (0.389$\pm$0.082 mJy)
in the residual map coincides with the lensed arc, and is consistent with
the difference between the integrated and the peak flux in the
original continuum map ($\sim$0.4 mJy). We therefore adopt
$S_\nu$ = 0.39$\pm$0.08 mJy as the 2\,mm continuum emission toward
the background galaxy.

@@

The emission in the IRAC images is slightly extended. We thus use the
{\it HST} image ($\sim$0\farcs07 resolution) to determine
origins of their centroids, all of which are found to be
centered at the position corresponding to the lensed emission from the
background galaxy. To recover the diffuse background emission, we subtract a
point source model centered on the lensing galaxy, using the average
FWHM found by fitting a Gaussian profile to several field stars
with the \ncode{imexam} routine of IRAF.
We perform aperture photometry on the residual image
to obtain decomposed flux measurements from the background galaxy.
The photometries for the foreground galaxy are then obtained
by subtracting the background emission from the
observed total flux. The resulting photometry in
\Tab{photometry} are obtained after performing aperture correction
described in the IRAC Instrument Handbook\footnote{http://irsa.ipac.caltech.edu/data/SPITZER/docs/irac/iracinstrumenthandbook/} to
correct for the fact that the imaging was calibrated
using a 12" aperture, which is bigger than the aperture (5\farcs8) we used to
perform aperture photometry.
We then fit a power-law spectrum $S_\nu \propto \nu^\alpha$ to the
decomposed IRAC photometry to disentangle the observed total flux
at MIPS 24\micron. We find a spectral index of $\alpha\sim-$1.8 and
$\alpha\sim-$0.85 for the lensing galaxy and RXJ1131, respectively.
This is consistent with the mean 3.6$-$8\micron spectral slope of
$\alpha$\,=\,$-$1.07$\pm$0.53 found for unobscured AGN
\citep{Stern05a}. An extrapolation of the fit to 24\micron
yields 33.96$\pm$0.01\,mJy and 25.19\,$\pm$\,0.03\,mJy
for the foreground galaxy and RXJ1131, respectively.
We note that the IRAC photometry includes
contributions from old stellar population and is prone to
dust extinction. Hence, the decomposed fluxes are only
our best estimate of the warm dust emission.
We incorporate the decomposed 24\micron point in our
SED fitting to provide some constraints on
the Wien tail beyond the dust peak
of the spectral energy distribution (SED) of RXJ1131.
Details of the SED modeling is presented in \Sec{SED}.

@@

We attempt to characterize the molecular gas kinematics using an empirically motivated disk model \citep[\eg][]{Courteau97a,Puech08a,Miller11a}:
\begin{equation}
V = V_0 + \frac{2}{\pi} V_{a} \arctan(\frac{R}{R_{t}}),
\end{equation}
where $V$ is the observed velocity, $V_0$ is the velocity at dynamical center,
$V_{a}$ is the asymptotic velocity, and $R_{t}$ is the ``turnover''
radius at which the rotation curve becomes flat.
We perform non-linear least square fitting using an orthogonal distance
regression to find the best-fit parameters,
taking into account the uncertainties in both velocity (channel width) and
distance offset. We also place an upper limit on $R_{t}$\,$<$15 kpc
to keep this parameter physical \citep[\eg][]{Puech08a,Miller11a}.
The parameter uncertainties are inferred based on Monte Carlo simulation
of 500 iterations, where the input parameters are perturbed
according to random Gaussian distributions of sigmas
corresponding to their uncertainties.
Using this model, we find $V_{a}$\,=\,975\,$\pm$\,387 \kms,
$R_{t}$\,=\,10.6\,$\pm$\,5.7\,kpc, and $V_0$\,=\,28\,$\pm$\,40 \kms.
We note, however, that since emission is not detected beyond the flat regime
of the rotation curve, the asymptotic velocity is poorly constrained and
the ``turnover'' radius is at most a lower limit.
In particular, $V_{a}$ and $R_{t}$ are highly correlated with a
Pearson coefficient $R$\,=\,0.998, and 0.027 between $V_{a}$ and $V_0$.

@@

\subsection{SED modeling} \label{sec:SED}
We fit dust SED models to the
24\micron$-$2.2\,mm photometry in \Tab{photometry}, where we also
include the IRAS 60\micron and 100\micron upper limits
to constrain the dust peak.
The fit is performed with the code
\ncode{mbb\_emcee} \citep[\eg][]{Dowell14a}, which
uses instrumental response curves to perform color correction on-the-fly.
The first SED model is a modified-blackbody
function with a power-law attached to the
blue/Wien side to account for an excess in the MIR owing to warm dust emission.
%We add $\sim$15\% calibration uncertainties in quadrature to obtain the total
%uncertainties for the PdBI continuum in our fitting procedure.

We find a power-law slope of BLAH,

% describe mbb_emcee
We check for chain convergence by requiring the autocorrelation
length for each model parameter is less than the number of steps
taken during the burn-in phase (which are then discarded).
% parameter estimation problem
For these parameter estimation problems using MCMC to sample,
the full answer is the full joint posterior PDF.
We here report the summary statistics (global mode (where min. chi),
marginal mode (in log file),
posterior median (instead of posterior means; need to compute) and the
1$\sigma$ confidence intervals in the marginal PDF) to reflect the PDF shape.

% upper limit using decomposed 24 um
% lower limit using fit without 24 um, but do I still want to fit a power-law for this?
The best-fit SEDs are shown in \Fig{SED} and the resulting parameters
are listed in \Tab{SED}.
Our SED model find a rest-frame characteristic dust temperature of BLAH
(cf. 60$-$70\,K in the cores of local ULIRGs/SB; Pauzzo+11, Rangwala+10),
\fir (rest-frame 42.5-122.5 \micron) of blah, and a dust mass of blah.
For the mass absorption coefficient, we adopt BLAH (rest frame; Dunne et al. 2003).
Previous studies find that in the absence of mid-IR data, a powerlaw slope of $\alpha$ = 2 is consistent with SB and slightly more shallow, $\sim$1.5, for SB/AGN composite systems (Blain et al., 2003; Casey, 2012; Koss et al., 2013).

We also fit an SED model excluding the decomposed 24.0\micron data.

\begin{figure*}[tbph]
\centering
\includegraphics[width=0.8\textwidth]{../Figures/FullSED.eps}
\caption{
WISE flux is slightly higher than the IRAC point, probably due to a smaller
aperture used in IRAC extraction, both of which were taken from the archive.
Spitzer IRAC:
Aperture flux in 5\farcs diameter with aperture correction applied.
Spitzer/MIPS: PSF fit flux, as source size <
native PSF FWHM of 6".
Might want to add NVSS constraints.
The data are not corrected for dust extinction.
\label{fig:SED}}
\end{figure*}

\subsubsection{SB versus AGN contribution}
% brief argument
What does it imply to the dust properties if we see the QSO.. contribution to the hot dust is not as much as if we were viewing it edge-on? --> warm dust and cold dust.

We quantify the AGN contribution to total IR.
important how the power source differ from high-z or local.

To derive the intrinsic properties, we correct for lensing using the
median magnification factor derived from the CO lens models.
Assuming a Salpeter initial mass function, we find the SFR of blah using the
lensing-corrected \fir luminosity.

\subsection{Gas properties}
In this section, we derive the gas properties of RXJ1131 based on \bco,
and compare the derived quantities against those reported by
\citet[hereafter C13]{Combes13a}, which is the largest sample of
ULIRGs at intermediate redshift (0.6\,$<$\,$z$\,$<$\,1.0) with CO measurements,
to place RXJ1131 in context of existing studies. Their results are
based on \bco and \rot{4}{3} line observations on the IRAM 30-m dish.

>>>>

\citestyle{aa}
\shorttitle{Molecular Gas in the Lensed Wet-Merger RXJ1131 at $z$\,=\,0.65}
\shortauthors{Leung \& Riechers}

@@

% moment 0 map: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth, first and second moment maps shown in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube, over $\Delta v$$\sim$750 \kms.
The first and second moment maps are clipped at 3$\sigma$.
The peak flux density is 8.12 $\pm$ 0.30 Jy\,\kms\,beam\pmOne in the intensity-integrated
map. The deconvolved source size is estimated to be 5\farcs1$\pm$0\farcs72 $\times$ 3\farcs72$\pm$0\farcs66
and thus the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as 2D-Gaussian,
the fit recovers the line intensity enclosed within the emitting
region, we therefore take this as an estimate on the extent of the lensed
emission. On the other hand, if we assume the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein Ring'' and
the lensed knots (see \Fig{CO21mom}).

@@

The emission of the red component coincides with the Einstein Ring seen in the
optical image, with most of its apparent flux originating from the lensed arc
in the Southeast, whereas the blue component is predominately coming from
solely the lensed arc shown in \Fig{CO21mom}. To illustrate this, we show the
channel maps of 21.5\,\kms width and spatial spectra of 1\farcs5 resolution in
\Fig{chanmap} and \Fig{spatialSpec}, respectively. The figures
show that emission
is present to the West, peaking toward the lensing arc (black across in
\Fig{chanmap}) in the red wing, and shifts to the East with decreasing velocity
(blue wing).
This indicates extended emission in the source plane, where
emission from different kinematic components is separated by different amounts
relative to the caustics. One consequence of such is the observed asymmetric
line profile. A detailed analysis on differential lensing of \bco is presented
in \Sec{differential}.

@@

\subsection{Origin of the \bco Emission} \label{sec:origin} %DONE
Evidently, the CO emission is co-spatial with the lensed optical emission in
the background AGN host galaxy.
The question is then whether the optically
faint companion also contributes to the CO flux and thus contains a
non-negligible amount of molecular gas.
Indeed, based on our lens model presented in \Sec{lensmodel}, we find it
highly likely that part of the CO emission originates from
the companion. We thus classify RXJ1131 as a ``wet-wet'' merger.
However, we cannot quantify this with a mass ratio given the
spatial resolution, nor can we rule out the possibility that
the ``companion'' is instead a gas component/tidal feature ripped out
from previous interaction.
In the subsequent sections, we will interpret
the this additional source component as gas emission
in a nearby galaxy (i.e. a ``wet-wet'' merger), as suggested by previous studies of RXJ1131 (C06, B08).

@@

We detect PdBI 2\,mm continuum in \Fig{cont}. The integrated flux density is
1.2$\pm$0.2 mJy, with a peak flux
$S_\nu$\,=\,800$\pm$88\,\micron Jy\,beam\pmOne
centering on the lensing galaxy. A slightly extended emission is also seen
along the lensed arc. This suggests that the detected emission comes from
both the foreground galaxy and the background galaxy and that the
emission is marginally resolved along its major axis.
We subtract a point source model in $uv$-plane to remove the unresolved
emission toward the foreground galaxy. The peak flux (0.39\,$\pm$\,0.08\,mJy)
in the residual map coincides with the lensed arc, and is consistent with
the difference between the integrated and the peak flux in the
original continuum map ($\sim$0.4 mJy). We therefore adopt
$S_\nu$ = 0.39$\pm$0.08 mJy as the 2\,mm continuum emission toward
the background galaxy.

@@

The emission in the IRAC images is slightly extended. We thus use the
{\it HST} image ($\sim$0\farcs07 resolution) to determine
origins of their centroids, all of which are found to be
centered at the position corresponding to the lensed emission from the
background galaxy. To recover the diffuse background emission, we subtract a
point source model centered on the lensing galaxy, using the average
FWHM found by fitting a Gaussian profile to several field stars
with the \ncode{imexam} routine of IRAF.
We perform aperture photometry on the residual image
to obtain decomposed flux measurements from the background galaxy.
The photometries for the foreground galaxy are then obtained
by subtracting the background emission from the
observed total flux. The resulting photometry in
\Tab{photometry} are obtained after performing aperture correction
described in the IRAC Instrument Handbook\footnote{http://irsa.ipac.caltech.edu/data/SPITZER/docs/irac/iracinstrumenthandbook/} to
correct for the fact that the imaging was calibrated
using a 12" aperture, which is bigger than the aperture (5\farcs8) we used to
perform aperture photometry.
We then fit a power-law spectrum $S_\nu \propto \nu^\alpha$ to the
decomposed IRAC photometry to disentangle the observed total flux
at MIPS 24$\micron$. We find a spectral index of $\alpha\sim-$1.8 and
$\alpha\sim-$0.85 for the lensing galaxy and RXJ1131, respectively.
This is consistent with the mean 3.6$-$8\,$\micron$
spectral slope of
$\alpha$\,=\,$-$1.07\,$\pm$\,0.53 found for unobscured AGN
\citep{Stern05a}. An extrapolation of the fit to 24$\micron$
yields 33.96\,$\pm$\,0.01\,mJy and 25.19\,$\pm$\,0.03\,mJy
for the foreground galaxy and RXJ1131, respectively.
We note that the IRAC photometry includes
emission from old stellar population and is prone to
dust extinction. Hence, the decomposed fluxes are only
our best estimate of the warm dust emission.
We incorporate the decomposed 24$\micron$ point in our
SED fitting to provide some constraints on
the Wien tail beyond the dust peak
of the spectral energy distribution (SED) of RXJ1131.
Details of the SED modeling is presented in \Sec{SED}.

@@

We attempt to characterize the molecular gas kinematics using an
empirically-motivated disk model \citep[\eg][]{Courteau97a,Puech08a,Miller11a}:
\begin{equation}
V = V_0 + \frac{2}{\pi} V_{a} \arctan(\frac{R}{R_{t}}),
\end{equation}
where $V$ is the observed velocity, $V_0$ is the velocity at dynamical center,
$V_{a}$ is the asymptotic velocity, and $R_{t}$ is the ``turnover''
radius at which the rotation curve becomes flat.
We perform non-linear least square fitting using an orthogonal distance
regression to find the best-fit parameters,
taking into account the uncertainties in both velocity (channel width) and
distance offset. We also place an upper limit on $R_{t}$\,$<$15 kpc
to keep this parameter physical \citep[\eg][]{Puech08a,Miller11a}.
The parameter uncertainties are inferred based on Monte Carlo simulation
of 500 iterations, where the input parameters are perturbed
according to random Gaussian distributions of sigmas
corresponding to their uncertainties.
Using this model, we find $V_{a}$\,=\,975\,$\pm$\,387 \kms,
$R_{t}$\,=\,10.6\,$\pm$\,5.7\,kpc, and $V_0$\,=\,28\,$\pm$\,40 \kms.
We note, however, that since emission is not detected beyond the flat regime
of the rotation curve, the asymptotic velocity is poorly constrained and
the ``turnover'' radius is at most a lower limit.
In particular, $V_{a}$ and $R_{t}$ are highly correlated with a
Pearson coefficient $R$\,=\,0.998, and 0.027 between $V_{a}$ and $V_0$.

@@

\subsection{SED modeling} \label{sec:SED}
We fit dust SED models to the
24\micron$-$2.2\,mm photometry in \Fig{SED}, where we also
include the IRAS 60$\micron$ and 100$\micron$ upper limits
to constrain the dust peak.
The fit is performed with the code
\ncode{mbb\_emcee} \citep[\eg][]{Dowell14a}, which samples the posterior
using an MCMC approach and uses instrumental
response curves to perform color correction on-the-fly.
The SED model consists of a modified-blackbody
function with a power-law attached to the
Wien side to account for an excess in the MIR owing to warm,
small dust emission.
% Previous studies find that in the absence of mid-IR data, a powerlaw slope
% of $\alpha$ = 2 is consistent with SB and slightly more shallow, $\sim$1.5, % for SB/AGN composite systems (Blain et al., 2003; Casey, 2012; Koss et al., 2013).
%
%We add $\sim$15\% calibration uncertainties in quadrature to obtain the total
%uncertainties for the PdBI continuum in our fitting procedure.
The model is thus described by five free parameters: the rest-frame characteristic dust
temperature (T$_{d}$), the emissivity index ($\beta$), the power-law index
($\alpha$), the flux normalization at 500$\micron$ ($f_{\rm norm}$), and
the observed-frame wavelength at which the emission
becomes optically thick ($\lambda_{0}$). We impose
an upper limit of 100\,K on $T_d$, a Gaussian prior centered around
$\mu$\,=\,1.9 with $\sigma$\,=\,0.3 on $\beta$, and an upper limit of
1000$\micron$ on $\lambda_0$.
We check for chain convergence by requiring the autocorrelation
length of each parameter to be less than the number of steps
taken for the burn-in phase (which are then discarded).
Here we report the modes %in log file
% summary statistics (global mode (where min. chi),
% marginal mode (in log file), posterior median (instead of posterior means;
% need to compute)
and the 1$\sigma$ confidence intervals in the marginal PDFs
as the best-fit parameters, as listed in \Tab{SED}.

\input{table_mbb}

In the first model, we include the 24$\micron$ data
to constrain the power-law index. Based on the
best-fit of this model, we find a
\fir luminosity (rest-frame 42.5-122.5\,$\micron$) of
3.81\petm{2.04}{-1.92}\E{12}\,\Lsun and a
dust mass of 22\petm{5}{-18}\E{8}\,\Msun.
For the mass absorption coefficient, we adopt
$\kappa$\,=\,2.64\,m$^2$kg\pmOne at 125.0$\micron$
(rest frame; \citealt{Dunne03a}).
The dust mass uncertainty does not
include those in the absorption coefficient.
These values are not corrected for lensing.

A fit including the MIR 24$\micron$ photometry
is likely an upper limit on the \fir luminosity arising from starburst.
% since the fit include contribution from the AGN (through power-law to MIR)
If we instead fit for a model excluding the 24$\micron$ constraint,
two major consequences are immediately apparent.
First, the power-law index is poorly-constrained (see \Tab{SED}).
Second, the steep power-law implies a small contribution
from the power-law regime
to the total IR luminosity as compared to the graybody.
Thus, the \fir luminosity in
this model should, in principle, corresponds to a
lower limit on the cold dust emission.
Using the best-fit parameters
for this model, we find a  \LFIR of 4.72\petm{2.54}{-2.26}\E{12}\,\Lsun and a $M_{\rm dust}$ of 11\petm{5}{6}\E{8}\,\Msun.
Again, the inferred quantities are uncorrected for lensing.
%The best-fit SEDs are shown in \Fig{SED} and the
% resulting parameters are listed in \Tab{SED}.

The dust temperature from both models is consistent with that
of ULIRGs at 0.6\,$<$\,$z$\,$<$\,1.0 (54\,$\pm$\,5\,K, \citealt{Combes13a}).
We note the \fir luminosities from both models are comparable, which is
not surprising given the lack of constraints in the MIR. In fact, the
\fir luminosity from the latter fit is slightly higher than the former, but
consistent within their uncertainties.
For the subsequent analysis, we adopt the physical quantities
from the former model. The choice of which fit does not affect
the derived star formation rate (SFR) given the similar \LFIR.
Yet, the dust mass is higher in the former model (but consistent within the
uncertainties), and thus the gas-to-dust ratio in \Sec{properties} 
should be considered as an estimate.

\begin{figure*}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{../Figures/FullSED.eps}
\caption{SED model fits to the dust photometry of RXJ1131.
WISE flux is slightly higher than the IRAC point, probably due to a smaller
aperture used in IRAC extraction, both of which were taken from the archive.
Spitzer IRAC:
Aperture flux in 5\farcs diameter with aperture correction applied.
Spitzer/MIPS: PSF fit flux, as source size <
native PSF FWHM of 6".
Might want to add NVSS constraints.
The data are not corrected for dust extinction.
\label{fig:SED}}
\end{figure*}

We correct for lensing using the median magnification factor 
derived from the CO lens models. Assuming a Salpeter initial 
mass function, we find the SFR of blah using a standard SK law.

\subsection{Gas properties} \label{sec:properties}
In this section, we derive the gas properties of RXJ1131 based on \bco,
and compare the derived quantities against those reported by
\citet[hereafter C13]{Combes13a}, which is the largest sample of
ULIRGs at intermediate redshift (0.6\,$<$\,$z$\,$<$\,1.0) with CO measurements,
to place RXJ1131 in context of existing studies. Their results are
based on \bco and \rot{4}{3} line observations on the IRAM 30-m dish.
